---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(dplyr)
library(openf1r)

if (FALSE) {
# Check for functions which do not have an @examples block for roxygen
system("grep -c examples man/*Rd", intern = TRUE) |> 
  grep(":0$", x = _, value = TRUE)
}

if (FALSE) {
  covr::report(covr::package_coverage(
    line_exclusions = list()
  ))
}

if (FALSE) {
  pkgdown::build_site(override = list(destination = "../coolbutuseless.github.io/package/openf1r"))
}

# Makevars options to do some deep testing for CRAN

# Type conversions are sane
# PKG_FLAG=-Wconversion

# Pointer overflow checks i.e. dodgy pointer arithmetic
# PKG_CFLAGS+=-fsanitize=pointer-overflow -fsanitize-trap=pointer-overflow
# Then run in the debugger:
# R -d lldb 
# run
# testthat::test_local()

```

# openf1r

<!-- badges: start -->
![](https://img.shields.io/badge/cool-useless-green.svg)
[![CRAN](https://www.r-pkg.org/badges/version/openf1r)](https://CRAN.R-project.org/package=openf1r)
[![R-CMD-check](https://github.com/coolbutuseless/openf1r/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/coolbutuseless/openf1r/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

`{openf1r}` is a package for fetching Formula 1 race data from the 
[OpenF1 API](https://openf1.org).


### What's in the box


* `f1_cars()`	Some data about each car, at a sample rate of about 3.7 Hz.
* `f1_drivers()`	Provides information about drivers for each session.
* `f1_intervals()`	Fetches real-time interval data between drivers and their gap to the race leader. Available during races only, with updates approximately every 4 seconds.
* `f1_laps()`	Provides detailed information about individual laps.
* `f1_locations()`	The approximate location of the cars on the circuit, at a sample rate of about 3.7 Hz.
* `f1_meetings()`	Provides information about meetings i.e. a Grand Prix or testing weekend. Usually includes multiple sessions (practice, qualifying, race, ...).
* `f1_pits()`	Provides information about cars going through the pit lane.
* `f1_positions()`	Provides driver positions throughout a session, including initial placement and subsequent changes.
* `f1_race_control()`	Provides information about race control (racing incidents, flags, safety car, ...).
* `f1_radios()`	Provides a collection of radio exchanges between Formula 1 drivers and their respective teams during sessions.
* `f1_sessions()`	Provides information about sessions - i.e. a distinct period of track activity during a Grand Prix (practice, qualifying, sprint, race, ...).
* `f1_stints()`	Provides information about individual stints i.e. a period of continuous driving by a driver during a session.
* `f1_weather()`	The weather over the track, updated every minute.
* `clear_cache()`	Clear cached downloads
* `segments_sector()`	Information about the codes for mini-segments

## Installation

<!-- This package can be installed from CRAN -->

<!-- ``` r -->
<!-- install.packages('openf1r') -->
<!-- ``` -->

You can install the latest development version from
[GitHub](https://github.com/coolbutuseless/openf1r) with:

``` r
# install.package('remotes')
remotes::install_github('coolbutuseless/openf1r')
```

<!-- Pre-built source/binary versions can also be installed from -->
<!-- [R-universe](https://r-universe.dev) -->

<!-- ``` r -->
<!-- install.packages('openf1r', repos = c('https://coolbutuseless.r-universe.dev', 'https://cloud.r-project.org')) -->
<!-- ``` -->

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(dplyr)
library(ggplot2)
library(openf1r)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Find the session info for the 2025 Australian GP Race
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
f1_sessions(year = 2025, country_code = 'AUS') %>%
  select(1:6, session_key) %>%
  knitr::kable()

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Which drivers were there?
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
f1_drivers(session_key = 9693) %>%
  select(1:5, name_acronym) %>%
  knitr::kable()

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# How did Piastri do (driver #81) ?
# Race position data indicates Piastri started 2nd but finished 9th
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
pos <- f1_positions(session_key = 9693, driver_number = 81) %>%
  mutate(date = lubridate::ymd_hms(date))

ggplot(pos) +
  geom_line(aes(date, position)) + 
  geom_point(aes(date, position)) +
  geom_text(aes(date, position, label = position), hjust = -0.3, vjust = -0.3) +
  theme_bw() + 
  scale_y_reverse(limits = c(20, 0)) +
  labs(
    title = "Oscar Piastri's race position",
    subtitle = "2025 Australian F1 Grand Prix"
  )
  
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# What does the lap position data look like?
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
loc <- f1_locations(session_key = 9693, driver_number = 81)
ggplot(loc) + 
  geom_path(aes(x, y), alpha = 0.2) + 
  theme_bw() + 
  coord_equal() +
  labs(
    title = "Oscar Piastri's race track location",
    subtitle = "2025 Australian F1 Grand Prix"
  )


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Race incidents
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
race_control <- f1_race_control(session_key = 9693) %>%
  filter(flag %in% c('YELLOW', 'DOUBLE YELLOW', 'RED')) %>%
  select(date, flag, lap_number, message, sector) %>%
  mutate(date = lubridate::ymd_hms(date))

race_control %>%
  knitr::kable()


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Lap speeds - including outlaps
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
laps <- f1_laps(session_key = 9693, driver_number = 81)
ggplot(laps) + 
  geom_line(aes(lap_number, lap_duration)) +
  scale_y_continuous(limits = c(0, NA)) +
  theme_bw() +
  labs(
    title = "Oscar Piastri's race track location\nYellow lines indicate yellow flags",
    subtitle = "2025 Australian F1 Grand Prix"
  )


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Race incidents superimposed on Piastri's lap times
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
ggplot(laps) + 
  geom_line(aes(lap_number, lap_duration)) +
  geom_vline(data = race_control, aes(xintercept = lap_number), color = 'orange') +
  scale_y_continuous(limits = c(0, NA)) +
  theme_bw() +
  labs(
    title = "Oscar Piastri's race track location",
    subtitle = "2025 Australian F1 Grand Prix"
  )


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# All drivers and their race position for this race
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
pos <- f1_positions(session_key = 9693) %>%
  mutate(
    date = lubridate::ymd_hms(date),
    position = as.factor(position),
    driver_number = as.factor(driver_number)
  )

drivers <- f1_drivers(session_key = 9693) %>%
  mutate(driver_number = as.factor(driver_number))

pos_final <- pos %>%
  group_by(driver_number) %>%
  filter(row_number() == n()) %>%
  ungroup() %>%
  left_join(drivers, by = 'driver_number')


ggplot(pos) + 
  geom_path(aes(date, position, group = driver_number, color = driver_number)) + 
  geom_text(data = pos_final, aes(date, position, label = full_name, color = driver_number),
            hjust = -0.3) +
  theme_bw() + 
  labs(
    title = "Race position of all drivers",
    subtitle = "2025 Australian F1 Grand Prix"
  ) + 
  theme(legend.position = 'none') +
  expand_limits(x = max(pos_final$date) + 3000)
```





## Related Software

* [f1dataR](https://cran.r-project.org/package=f1dataR)


